Dependent Object Types

### 2012-06-30 doc/semester/report.tex
functions as objects

\begin{align*}
\mlnew u \ldots {
\ \mlapp{\abs x \Top x}{
\ \ \mlapp{\abs f {S \tfun U} f}{
\ \ \ \mlapp{\abs f {S \tfun T} f}{
\ \ \ \ \mlapp{\abs f {S \tfun S} f}{
\ \ \  \ \ \abs x S x}}}}}
\end{align*}

@@

\subsection{Non-Expanding Types and Subtyping Transitivity}

@@

>>>>

\begin{align*}
&\mlnew u \ldots {
\ \mlapp{\abs x \Top x}{
\ \ \mlapp{\abs f {S \tfun U} f}{
\ \ \ \mlapp{\abs f {S \tfun T} f}{
\ \ \ \ \mlapp{\abs f {S \tfun S} f}{
\ \ \  \ \ \abs x S x}}}}}
\end{align*}

@@

\subsection{Non-Expanding Types and Subtyping Transitivity}\label{nonexp}

@@

\subsection{Narrowing}\label{narrowing}

\subsubsection{Functions as Objects}\label{fun}

The original DOT calculus has both object and function types. It
doesn't have built-in methods, but they can be expressed as functional
labels. On the other hand, Scala doesn't have a built-in function
type. Instead, it has both methods and fields, and encodes functions
as objects with an apply method. Here, I show that the DOT model as
originally designed is problematic.

A concrete object can be a subtype of a function type without a
function ever being defined. Consider:
\begin{align*}
&\mlnew u {\Top \refine z {\Ldecl C {\Top \tfun \Top} {\Top \tfun \Top}} \ldefs{}} {
\mlnew f {u.C \ldefs{}} {
\ldots
}}
\end{align*}

Now, $f$ is a subtype of $\Top \tfun \Top$, but $\app f {(\abs x \Top
  x)}$ is stuck (and, rightfully, doesn't typecheck). But we can use
narrowing to create a counterexample to preservation: $\app {(\abs g
  {\Top \tfun \Top} {g (\abs x \Top x)})} f$.
